\newacronym{sacd}{SACD}{snow cover area calculation and display}

\newacronym{psu}{psu}{practical salinity unit}

\newacronym{tp}{TP}{Tibetan Plateau}

\newacronym{qc}{QC}{quality control}

\newacronym{ims}{IMS}{Interactive Multisensor Snow and Ice Mapping System}

\newacronym{tsm}{TSM}{Tibetan Snow Man}

\newacronym{avhrr}{AVHRR}{Advanced Very High-Resolution Radiometer}

\newacronym{modis}{MODIS}{Moderate Resolution Imaging Spectrometer}

\newacronym{gms}{GMS}{Geostationary Meteorological Satellite}

\newacronym{metosat}{METEOSAT}{European Weather Satellite}

\newacronym{cmc}{CMC}{Common Mapping Client}

\newacronym{dat}{DAT}{NASA Sea Level Change Data Analysis Tool}

\newacronym{ssmi}{SSMI}{Special Sensor Microwave Imager}

\newacronym{dmsp}{DMSP}{Defense Meteorological Satellite Program}

\newacronym{amsu}{AMSU}{Advanced Microwave Sounding Unit}

\newacronym{nh}{NH}{Northern Hemisphere}

\newacronym{nsidc}{NSIDC}{National Snow and Ice Data Center}

\newacronym{laeap}{LAEAP}{Lambert azimuthal equal area projection}

\newacronym{foss}{FOSS}{free open source software}

\newacronym{t/s/p}{T/S/P}{temperature salinity pressure}

\newacronym{gdac}{GDAC}{Global Data Assembly Centres}

\newacronym{dac}{DAC}{Data Assembly Centres}

\newacronym{ftp}{FTP}{File Transfer Protocol}

\newacronym{loess}{LOESS}{local regression}

\newacronym{wmo}{WMO}{World Meteorological Organization}

\newacronym{aic}{AIC}{Akaike information criterion}

\newacronym{xml}{XML}{extensible markup language}

\newacronym{json}{JSON}{JavaScript object notation}

\newacronym{url}{URL}{Uniform Resource Locator}

\newacronym{api}{API}{application programming interface}


\newacronym{noaa}{NOAA}{National Oceanic and Atmospheric Administration}

\newacronym{nasa}{NASA}{National Aeronautics and Space Administration}

\newglossaryentry{filezilla}{
 name=FileZilla,
 description={Free software, cross-platform FTP application, consisting of FileZilla Client and FileZilla Server. Client binaries are available for Windows, Linux, and macOS.}}

\newglossaryentry{restfull}{
 name=RESTfull,
 description={REpresentational State Transfer, or RESTful, web services provide interoperability between computer systems on the Internet.}}

\newglossaryentry{aws}{
 name=AWS,
 description={Amazon Web Services (AWS) provides on-demand cloud computing platforms to individuals, companies and governments, on a paid subscription basis.}}
 
\newglossaryentry{hadoop}{
 name=Hadoop,
 description={Apache Hadoop is an open-source software framework used for distributed storage and processing of datasets of big data using the MapReduce programming model. It consists of computer clusters built from commodity hardware.}}

\newglossaryentry{rsync}{
 name=rsync,
 description={rsync is a utility for efficiently transferring and synchronizing files across computer systems, by checking the timestamp and size of files.}}

\newglossaryentry{leaflet}{
  name=leaflet,
  description={Open source JavaScript library used to build web mapping applications. Includes plugins for drawing objects, using different map projections, and using different map tiles.}}

\newglossaryentry{mvc}{
  name=MVC,
  description={Model–view–controller (MVC) is a software architectural pattern commonly used for developing user interfaces that divides an application into three interconnected parts.}}

\newglossaryentry{cron}{
  name=cron,
  description={A time-based job scheduler in Unix-like computer operating systems. Cron enables scripts to run at periodically at a fixed time.}}

\newglossaryentry{dataseries}{
 name=dataseries,
 description={A one-dimensional labeled array capable of holding any data type (integers, strings, floating point numbers, Python objects, etc.).}}

\newglossaryentry{dataframe}{
  name=DataFrame,
  description={2D Table like object with columns of potentially different types. Similar to a SQL table or MS Excel spreadsheet.}}


\newglossaryentry{dod}{
  name=document-oriented database,
  description={Is a type of NoSQL database, used to store semi-structured data, often in key-value schemas, such as XML or JSON. Unlike traditional relational databases, the data model in a document database is not structured in a table format of rows and columns. This allows flexibility in data structures and modeling.}}


\newglossaryentry{Iridium}{
  name=Iridium,
  description={Iridium Communications Inc. is a company that supplies Argo floats an antenna for data transmission.}}
\newglossaryentry{Ipython}{
  name=Ipython,
  description={IPython is an interactive command shell for the Python language.}}
  
\newglossaryentry{Jupyter}{
  name=Jupyter,
  description={A browser based notebook. Jupyter is a FOSS web application that allows, text, equations, and figure integration in a number of languages, includeing Python.}}
  
\newglossaryentry{express}{
  name=express.js,
  description={The standard server web application framework for Node.js. Used to make web-apps.}}

\newglossaryentry{node}{
  name=node.js,
  description={Node.js JavaScript run-time environment for executing JavaScript code server-side. Historically, javaScript was run on the client side only, with Node.js, javascript is run like any high-level interepreted language.}}
  
\newglossaryentry{mongodb}{
  name=MongoDB,
  description={MongoDB (from humongous) is document-oriented database program, classified as a NoSQL.  MongoDB uses JSON-like documents with schemas. MongoDB is developed by MongoDB Inc., and is published under a combination of the GNU Affero General Public License and the Apache License.}}

\newglossaryentry{WGS-84}{
  name=WGS-84,
  description={World Geodetic System established in 1984, is an Earth-centered and Earth-fixed terrestrial reference system and geodetic datum describing the Earth's shape, size, gravity and geomagnetic fields.}}

\newglossaryentry{.asc}{
  name=.asc,
  description={File that stores ascii characters}}

\newglossaryentry{HDF5}{
  name=HDF5,
  description={Hierarchical Data Format is a set of file formats designed to store and organize large amounts of data.}}

\newglossaryentry{NetCDF}{
  name=NetCDF,
  description={Network Common Data Form is a file storage system, denoted by .nc. It allows the creation, access, and sharing of array-oriented scientific data. The ARGO program releases data NetCDF on two FTP sites.}}
  
\newglossaryentry{ARGO}{
  name=ARGO,
  description={Array for Real-time Geostrophic Oceanography is a global array of 3,800 free-drifting profiling floats whos primary measurement includes temperature and salinity of the upper 2000 m of the ocean.}}